Syntactic realization with data-driven neural tree grammars

### 2016-10-10 paper/draft.tex
first round of wording changes

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

@@

% \author{Brian McMahan \\
%   Computer Science, Rutgers University \\
%   {\tt brian.mcmahan@rutgers.edu} \\\And
%   Matthew Stone \\
%   Computer Science, Rutgers University \\
%   {\tt matthew.stone@rutgers.edu} \\}

@@

With traditional stochastic modeling techniques, NLG researchers have
had to predict choices using factored models with handcrafted
representations and strong independence assumptions, in order to avoid
combinatorial explosions and address the sparsity of training data.
By contrast, in this paper, we leverage recent advances in deep
learning to develop new models for syntactic choice that free
engineers from many of these decisions, but still generalize more
effectively and match human choices more closely than traditional
techniques.

We adopt the characterization of syntactic choice from
\newcite{bangalore2000exploiting} which uses a stochastic tree model and a language model to produce a linearized string from an unordered,
unlabeled dependency graph.
%
The first step to producing a linearized string is
to assign each node an appropriate supertag---a fragment of a parse tree with
a leaf left open for a lexical item---using a stochastic tree model.
%A language model is then used to select among the many possible derivation t
The resulting assignments are consistent with many possible derivation trees
because of adjunct ambiguitity.
%
To finish the linearization, a language model is used to select among them.
%
While improving the language model would improve the linearized string,
this work focuses on more accurately predicting the correct supertags from
unlabeled dependency trees.

In this work, we make two novel contributions to improve upon the stochastic tree model:
a supertag embedding technique which uses convolutional neural networks and
a recurrent tree network which can model hierarchical relationships.
%
By embedding supertags and using a recurrent tree network for the relationships between them,
we enable more information to be used at each node.
%
We evaluate our contributions in two ways: first, by varying the technique used to embed
supertags, and then by comparing a feed-forward model against our
recurrent tree model.

Our presentation begins in \ref{sec:tree} with an introduction to tree
grammars and a deterministic metholodogy for inducing the elementary trees of the grammar.
%
Next, \ref{sec:neural} presents the techniques we have developed to represent
a tree grammar using a neural architecture.
%
Then, in \ref{sec:models}, we describe the specific model we have
implemented, and in \ref{sec:algo}, we describe the algorithms used to describe the classification results.
%
The experiments in \ref{sec:expt} demonstrate the improvement of the
model over baseline results based on previous work on stochastic
surface realization.
%
We conclude with a brief discussion of the future potential for neural
architectures to predict grammatical choices in NLG.

@@

Broadly, tree grammars are a family of tree rewriting formalisms that
modify structure to produce strings.  The atomic syntactic units are
called elementary trees; elementary trees combine using tree-rewrite
rules to form derived phrase structure trees describing complex
sentences.  Inducing a tree grammar involves fixing a formal inventory
of structures and operations for elementary trees and then inferring
instances of those structures to match corpus data.

@@

The canonical tree grammar is perhaps lexicalized tree-adjoining
grammar (LTAG) \cite{joshi1991tree}.
%
The elementary trees of LTAG consist of two disjoint sets with distinct operations
%%%
The elementary trees of LTAG are disjoint sets based on the operations
they can perform: initial trees can perform substitution operations
and auxiliary trees can perform adjunction operations.
%
The substitution operation replaces a non-terminal leaf of a target
tree with an identically-labeled root node of an initial tree.
%
The adjunction operation modifies the internal structure of a target
tree at by expanding a node identically-labeled with the root and a
distinguished foot note in the auxiliary tree.
%
The lexicalization of the the grammar requires each elementary tree to
have at least one lexical item as a leaf.

@@

We adopt one particular restricted adjunction, called
sister-adjunction or \emph{insertion}, which allows trees to attach to
an interior node and add itself as a first or last child
\cite{chiang2000statistical}.
%
Chiang's sister-adjunction allows for the flat structures in
the Penn Treebank while limiting the formalism to context-free power.

@@

We follow previous work in using bracketed phrase structure corpora
and deterministic rules to induce the grammar
\cite{chiang2000statistical,bangalore2001impact}.
%
Broadly, the methodology is to split the observed trees into the
constituents which make it up, according to the grammar formalism.
%
We use head rules
\cite{chiang2000statistical,collins1997three,magerman1995statistical}
to associate internal nodes in a bracketed tree with the lexical item
that owns it.
%
We use additional rules to classify some children as complements,
corresponding to substitution sites and root notes of complement
trees; and other children as adjuncts, corresponding to insertion
trees that combine with the parent node, either to the right or to the
left of the head; this allows us to segment the tree into units of
substitution and insertion.%
\footnote{One particular downside of deterministically constructing
  the grammar this way is that it can produce an excess of superfluous
  elementary trees.
%
We minimize this by collapsing repeated projections in the treebank.
%
Other work has provided Bayesian models for reducing grammar
complexity by forcing it to follow Dirichlet or Pitman-Yor processes
\cite{Cohn2010}---an interesting direction for future work.}

@@

%There is a tension between vectorized representations and dynamic, nonlinear data structures.
%
%The tension is emphasized by requirement of specialized hardware that both dramatically increases the data throughput and reduces the space of representable models\footnote{Specifically, to run on Graphical Processing Units (GPUs), model structure must be known at compile time or incur heavy computational costs.}.
%
This work makes two contributions to improve stochastic tree modeling with neural networks.
%This work makes two contributions to reduce that tension and progress methodologies for representing nonlinear data structures in linear spaces.
%
First, supertags are represented as vectors through convolutional coding,
opening up the recent repertoire of vector space techniques.
%
Second, hierarchical dependence between nodes are modeled using a recurrent tree
network.
%
%recurrent neural network variant that isn't restricted to conditioning hidden states on their temporal predecessors.
%In brief, this technique is made possible through recent advanced in StackLSTMs \cite{dyer2015transition} and the use of topological indexing of recurrent states.% hidden states of

@@

%The first step to embedding supertags with a convolutional neural network is to
%embed each of its nodes and combine them to form a tensor.
%
%
%To embed a supertag, we first embed all of its nodes to form a tensor and then use a
%To embed a supertag and summarize it into a vector, each of its nodes are
%first embedded, combined to form a tensor, and then
%We embed supertags by embed each of its nodes, combine
To embed a supertag, each of its nodes are embedded, grouped to form a tensor, and then summarized into a single vector using a series of convolutional neural networks.
%
Since the weights of a convolutional neural network are optimized during training, the data should determine which parts of the supertag are important.
%
%The purpose of convolutional coding is to allow the data to determine which facets of a supertag are relevant and useful.
%
The more usual alternative approach is to treat each supertag as the atomic
unit---essentially a statistical singleton---and map it either to a feature or
an embedding vector.
%
We evaluate our contribution and compare these two approaches to get a sense for whether convolutional coding robustly generalizes across supertags.

The goal of embedding supertags is to summarize the tree structure into a single vector.
%
Inside the tree structure, there are two separate pieces of information which need to be summarized:
the syntactic category of each node and the role each node plays in the tree grammar.
%
The root receives a special role---either substitution or insertion to represent
the supertag's tree operation.
%g
The remaining nodes either have the substitution point role or the spine role---they are along the spine from root to the lexical attachment point.

@@

Fergus-N is a stochastic tree model which uses local parent-child information as inputs to a feed-forward network.
%
Each parent-child pair are treated as independent of all other pairs.
%
The probability of the parent's supertag is computed using an embedding of the pair's lexical material and an embedding of the child's supertag.
%
During training, the negative log probability of the observed parent supertags is minimized for each parent-child pair.
%
This objective function is formally defined in Eq. \ref{eq:fergusn_obj} where $tag_p$ is the parent supertag, $tag_c$ is the child supertag, $lex_p$ is the parent's lexical material, and $lex_c$ is the child's lexical material.
%
Note that the probability of supertags for the leaves of the tree are computed
with respect to their parent's lexical material.

@@

To efficiently decode the supertag classifications, we implement an A* algorithm to incrementally select consistent supertag assignments.
%
At each step, the algorithm uses a priority queue to select subtrees based on their inside-outside scores.
%
The inside score is computed as the sum of the log probabilities of the supertags in the subtree.
%
The outside score is the sum of the best supertag for nodes outside the subtree, similar to \newcite{lewis2014improved}.
%
Once selected, the subtree is attached to the possible supertags of its parent that are both locally consistent and consistent among its already attached children.
%
These resulting subtrees are placed into the priority queue and the algorithm iterates to progress the search.
%
The search concludes either when a single complete tree has been found\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed}.

@@

>>>>

@@

\author{Brian McMahan \and Matthew Stone \\
   Computer Science, Rutgers University \\
   {\tt brian.mcmahan, matthew.stone@rutgers.edu}}

@@

With traditional stochastic modeling techniques, NLG researchers have
had to predict choices using factored models with handcrafted
representations and strong independence assumptions, in order to avoid
combinatorial explosions and address the sparsity of training data.
%
By contrast, in this paper, we leverage recent advances in deep
learning to develop new models for syntactic choice that free
engineers from many of these decisions, but still generalize more
effectively, match human choices more closely, and enable more
efficient computations than traditional techniques.

We adopt the characterization of syntactic choice from
\newcite{bangalore2000exploiting}, which uses a stochastic tree model
and a language model to produce a linearized string from an unordered,
unlabeled dependency graph.
%
The first step to producing a linearized string is to assign each item
an appropriate \emph{supertag}---a fragment of a parse tree with a
leaf left open for a lexical item---using a stochastic model.
%
This involves applying a learned model to make predictions for the
syntax of each item and then searching over the predictions to find a
consistent assignment for the entire sentence.
%
The resulting assignments allow for many possible surface realization
outputs because they can underdetermine the order and attachment of
adjuncts.
%
To finish the linearization, a language model is used to select the
most likely surface form from among the alternatives.
%
While improving the language model would improve the linearized string,
this work focuses on more accurately predicting the correct supertags from
unlabeled dependency trees.

Our work exploits deep learning to improve the model of supertag
assignment in two ways.
%
First, we analyze the use of embedding technques to generalize across
supertags.  
%
Neural networks offer a number of architectures that can cluster tree
fragments during training; such models learn to treat related structures
similarly, and we show that they improve supertag assignments.
%
Second, we analyze the use of tree recurrences to track hierarchical
relationships within the generation process.
%
Such networks can track more of the generation context than a simple
feed-forward model; as a side effect, they can simplify the problem of
computing consistent supertag assignments for an entire sentence.
%
We evaluate our contributions in two ways: first, by varying the
technique used to embed supertags, and then by comparing a
feed-forward model against our recurrent tree model.

Our presentation begins in \ref{sec:tree} with an introduction to tree
grammars and a deterministic metholodogy for inducing the elementary
trees of the grammar.
%
Next, \ref{sec:neural} presents the techniques we have developed to
represent a tree grammar using a neural architecture.
%
Then, in \ref{sec:models}, we describe the specific models we have
implemented and the algorithms used to exploit the models in NLG.
%
The experiments in \ref{sec:expt} demonstrate the improvement of the
model over baseline results based on previous work on stochastic
surface realization.
%
We conclude with a brief discussion of the future potential for neural
architectures to predict NLG choices.

@@

Broadly, tree grammars are a family of tree rewriting formalisms that
produce strings as a side effect of composing primitive
hierarchical structures.  The basic syntactic units are called
elementary trees; elementary trees combine using tree-rewrite rules to
form derived phrase structure trees describing complex sentences.
Inducing a tree grammar involves fixing a formal inventory of
structures and operations for elementary trees and then inferring
instances of those structures to match corpus data.

@@

The canonical tree grammar is perhaps lexicalized tree-adjoining
grammar (LTAG) \cite{joshi1991tree}.
%
The elementary trees of LTAG consist of two disjoint sets with
distinct operations: initial trees can perform substitution operations
and auxiliary trees can perform adjunction operations.
%
The substitution operation replaces a non-terminal leaf of a target
tree with an identically-labeled root node of an initial tree.
%
The adjunction operation modifies the internal structure of a target
tree by expanding a node identically-labeled with the root and a
distinguished foot note in the auxiliary tree.
%
The lexicalization of the the grammar requires each elementary tree to
have at least one lexical item as a leaf.

@@

We adopt one particular restriction on adjunction, called
sister-adjunction or \emph{insertion}, which allows trees to attach to
an interior node and add itself as a first or last child
\cite{chiang2000statistical}.
%
Chiang's sister-adjunction allows for the flat structures in
the Penn Treebank while limiting the formalism to context-free power.

@@

We follow previous work in using bracketed phrase structure corpora
and deterministic rules to induce the grammar
\cite{bangalore2001impact,chiang2000statistical}.
%
Broadly, the methodology is to split the observed trees into the
constituents which make it up, according to the grammar formalism.
%
We use head rules
\cite{chiang2000statistical,collins1997three,magerman1995statistical}
to associate internal nodes in a bracketed tree with the lexical item
that owns it.
%
We use additional rules to classify some children as complements,
corresponding to substitution sites and root notes of complement
trees; and other children as adjuncts, corresponding to insertion
trees that combine with the parent node, either to the right or to the
left of the head.  This allows us to segment the tree into units of
substitution and insertion.%
\footnote{One particular downside of deterministically constructing
  the grammar this way is that it can produce an excess of superfluous
  elementary trees.
%
We minimize this by collapsing repeated projections in the treebank.
%
Other work has provided Bayesian models for reducing grammar
complexity by forcing it to follow Dirichlet or Pitman-Yor processes
\cite{Cohn2010}---an interesting direction for future work.}

@@

%There is a tension between vectorized representations and dynamic, nonlinear data structures.
%
%The tension is emphasized by requirement of specialized hardware that both dramatically increases the data throughput and reduces the space of representable models\footnote{Specifically, to run on Graphical Processing Units (GPUs), model structure must be known at compile time or incur heavy computational costs.}.
%
This work makes two contributions to improve stochastic tree modeling with neural networks.
%This work makes two contributions to reduce that tension and progress methodologies for representing nonlinear data structures in linear spaces.
%
First, supertags are represented as vectors through embedding
techniques that enable to model to generalize naturally over complex, but
related structures.
%
Second, hierarchical dependence between choices is modeled using a recurrent tree
network that can capture long-distance influences as well as local ones.
%
%recurrent neural network variant that isn't restricted to conditioning hidden states on their temporal predecessors.
%In brief, this technique is made possible through recent advanced in StackLSTMs \cite{dyer2015transition} and the use of topological indexing of recurrent states.% hidden states of

@@

%The first step to embedding supertags with a convolutional neural network is to
%embed each of its nodes and combine them to form a tensor.
%
%
%To embed a supertag, we first embed all of its nodes to form a tensor and then use a
%To embed a supertag and summarize it into a vector, each of its nodes are
%first embedded, combined to form a tensor, and then
%We embed supertags by embed each of its nodes, combine
To embed a supertag, we embed each node, group the resulting vectors
to form a tensor, and then summarize the tensor into a single vector
using a series of convolutional neural networks.
%
Since the weights of a convolutional neural network are optimized
during training, the data should determine which parts of the supertag
are important.
%
%The purpose of convolutional coding is to allow the data to determine which facets of a supertag are relevant and useful.
%
The more usual alternative approach is to treat each supertag as the atomic
unit---essentially a statistical singleton---and map it either to a feature or
an embedding vector.
%
We evaluate our contribution and compare these two approaches to get a sense for whether convolutional coding robustly generalizes across supertags.

To explain our technique, we begin by observing that each elementary
tree is a complex structure with nodes labeled by category and
assigned a role that enables further tree operations.
%
The root node's role represents the overall action associated with
that elementary tree---either substitution or insertion.
%
The remaining nodes either have the substitution point role or the
spine role---they are along the spine from root to the lexical
attachment point, and thus provide targets for further insertion.

@@

Fergus-N is a stochastic tree model which uses local parent-child information as inputs to a feed-forward network.
%
Each parent-child pair is treated as independent of all others.
%
The probability of the parent's supertag is computed using an embedding of the pair's lexical material and an embedding of the child's supertag.
%
During training, the negative log probability of the observed parent supertags is minimized for each parent-child pair.
%
This objective function is formally defined in Eq. \ref{eq:fergusn_obj} where $tag_p$ is the parent supertag, $tag_c$ is the child supertag, $lex_p$ is the parent's lexical material, and $lex_c$ is the child's lexical material.
%
Note that the probability of supertags for the leaves of the tree are computed
with respect to their parent's lexical material.

@@

To efficiently decode the supertag classifications, we implement an A* algorithm to incrementally select consistent supertag assignments.
%
At each step, the algorithm uses a priority queue to select subtrees based on their inside-outside scores.
%
The inside score is computed as the sum of the log probabilities of the supertags in the subtree.
%
The outside score is the sum of the best supertag for nodes outside the subtree, similar to \newcite{lewis2014improved}.
%
Once selected, the subtree is attached to the possible supertags of its parent that are both locally consistent and consistent among its already attached children.
%
These resulting subtrees are placed into the priority queue and the algorithm iterates to progress the search.
%
The search concludes either when a single complete tree has been found.\footnote{Although, the data has some noise so that sometimes there is no complete tree that can possibly be formed}

@@

\section*{Acknowledgments}

This research was supported in part by NSF IIS-1526723 and by a
sabbatical leave from Rutgers to Stone.
